---
title: "HW2"
author: "Yongzi Yu"
output: html_document
---

I'm an R Markdown document! 

# Problem1

```{r}
library(tidyverse)
library(tidyr)
library(readxl)
library(Hmisc)
library(dplyr)


trash_data= 
  read_excel("./data/trash.xlsx",skip=1,sheet=1)%>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered)%>%
  drop_na(dumpster)%>%
  mutate(sports_balls=round(sports_balls))
trash_data

pre_2019=read_excel("./data/trash.xlsx",skip=1,sheet=4)%>%
  janitor::clean_names() %>%
  drop_na()%>%
  mutate(year=2019)

pre_2018=read_excel("./data/trash.xlsx",skip=1,sheet=5)%>%
  janitor::clean_names() %>%
  drop_na()%>%
  mutate(year=2018)

pre_1819=full_join(pre_2018,pre_2019) %>%
  mutate(month=month.name[month])
pre_1819
sum(pre_1819$total)
```

- For Mr.Trash Wheel Sheet, there are `r nrow(trash_data)` rows and `r ncol(trash_data)` columns.    Key variables are `r colnames(trash_data)`.  
- For precipitation, there are `r nrow(pre_1819)` rows and `r ncol(pre_1819)` columns.  
Key variables are `r colnames(trash_data)`.  Total precipitation in 2018 is `r sum(pull(pre_2018,total))`.  Median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trash_data,year=="2017"),sports_balls))`.

# Problem2
```{r}
pols= 
  read_csv("./data/pols_month.csv") %>%
  separate(mon, c("year", "month","day"), sep="-") %>% #separate value
  mutate(month=month.name[(as.double(month))]) %>% #change month
  mutate(president=prez_gop+prez_dem) %>% #add variable
  select(-prez_gop,-prez_dem,-day) %>% #remove col 
  mutate(key=paste(as.character(year),as.character(month)))


snp=
  read_csv("./data/snp.csv")%>%
  separate(date, c("month","day","year"), sep="/") %>% #separate value
  mutate(month=month.name[(as.double(month))]) %>% #change month
  arrange(year,month)%>%
mutate(key=paste(as.character(year),as.character(month)))

unemployment=
  read_csv("./data/unemployment.csv")%>%
  pivot_longer(
    Jan:Dec,
    names_to = "month", 
    values_to = "rate")%>%
mutate(key=paste(as.character(Year),as.character(month)))

#join
pols_snp=left_join(pols,snp,by='key')
pols_snp_unempl=left_join(pols_snp,unemployment,by='key')
```

- there are `r nrow(pols_snp_unempl)` rows and `r ncol(pols_snp_unempl)` columns.    Key variables are `r colnames(pols_snp_unempl)`.  

# Problem 3

```{r}

names= 
  read_csv("./data/Popular_Baby_Names.csv")

#remove duplicates
names=
  names[!duplicated(names), ]
names %>%
  janitor::clean_names() %>%
  mutate(childs_first_name=capitalize(tolower(childs_first_name)))

#draw plot
names = 
  names %>% 
  janitor::clean_names() %>% 
  filter(year_of_birth=='2016') %>% 
  filter(gender=='MALE') %>% 
  filter(ethnicity == 'WHITE NON HISPANIC')


ggplot(names,aes(y=rank,x=count))+geom_point()
```

