---
title: "HW2"
author: "Yongzi Yu"
output: html_document
---

I'm an R Markdown document! 

# Problem1

```{r}
library(tidyverse)
library(tidyr)
library(readxl)

trash_data= 
  read_excel("./data/trash.xlsx",skip=1,sheet=1)%>%
  janitor::clean_names() %>%
  select(dumpster:homes_powered)%>%
  drop_na(dumpster)%>%
  mutate(sports_balls=round(sports_balls))
trash_data

pre_2019=read_excel("./data/trash.xlsx",skip=1,sheet=4)%>%
  janitor::clean_names() %>%
  drop_na()%>%
  mutate(year=2019)

pre_2018=read_excel("./data/trash.xlsx",skip=1,sheet=5)%>%
  janitor::clean_names() %>%
  drop_na()%>%
  mutate(year=2018)

pre_1819=full_join(pre_2018,pre_2019) %>%
  mutate(month=month.name[month])
pre_1819
sum(pre_1819$total)
```

- For Mr.Trash Wheel Sheet, there are `r nrow(trash_data)` rows and `r ncol(trash_data)` columns.    Key variables are `r colnames(trash_data)`.  
- For precipitation, there are `r nrow(pre_1819)` rows and `r ncol(pre_1819)` columns.  
Key variables are `r colnames(trash_data)`.  Total precipitation in 2018 is `r sum(pull(pre_2018,total))`.  Median number of sports balls in a dumpster in 2017 is `r median(pull(filter(trash_data,year=="2017"),sports_balls))`.

# Problem2
```{r}
pols_month= 
  read_csv("./data/pols_month.csv") %>%
  separate(mon, c("year", "month","day"), sep="-") %>%
  mutate(month=month.abb[(as.double(month))]) %>%
  mutate(president=prez_gop+prez_dem) %>%
  select(-prez_gop,-prez_dem,-day)
pols_month
```